---
title: "BMP Tcell"
author: "A.DeMartin"
date: "2024-02-21"
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## load packages
```{r load packages, warning=FALSE, include=FALSE}
library(ExploreSCdataSeurat3)
library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(org.Hs.eg.db)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(NCmisc)
```

## load file
```{r load merged file}
##load merged file 
fileNam <- "/Users/immbio/Desktop/Project/Angelina/HumanHeartCarTrans/data/Human_heart_allmerged_seurat.rds"
seuratM <- readRDS(fileNam)
table(seuratM$dataset)
table(seuratM$RNA_snn_res.0.25)
table(seuratM$orig.ident)
table(seuratM$clusterName)
```
##set color vectors 
```{r set color vector}
colclusterName <- c("#67001f", "#D53E4F", "#f4a582", "#FEE08B", "#003c30","#01665e","#66C2A5", "#3288BD","#BEAED4", "#c7eae5","#355C7D","#202547","#B45B5C","#8c510a")
names(colclusterName) <- c("CM","Fb","PerivFb","VSMC","Int","BEC1", "BEC2","LEC","NC","AdipoC","Mph","Mph2","Tcell","Tcell2")

coldiseaseCond <- c("#dfc27d","#BE3144","#202547","#355C7D","#779d8d")
names(coldiseaseCond) <- c("healthy", "explant", "visit1", "visit2", "visit3")
```


## BMP features
```{r BMP features}
FeaturePlot(seuratM, features = "ENSG00000125378.BMP4", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)
FeaturePlot(seuratM, features = "ENSG00000112175.BMP5", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)
FeaturePlot(seuratM, features = "ENSG00000125845.BMP2", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)
FeaturePlot(seuratM, features = "ENSG00000153162.BMP6", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)
FeaturePlot(seuratM, features = "ENSG00000168487.BMP1", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)
FeaturePlot(seuratM, features = "ENSG00000183682.BMP8A", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)
FeaturePlot(seuratM, features = "ENSG00000138756.BMP2K", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)

FeaturePlot(seuratM, features = "ENSG00000107779.BMPR1A", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)
FeaturePlot(seuratM, features = "ENSG00000204217.BMPR2", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)

FeaturePlot(seuratM, features = "ENSG00000166923.GREM1", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)
FeaturePlot(seuratM, features = "ENSG00000180875.GREM2", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)

FeaturePlot(seuratM, features = "ENSG00000019991.HGF", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)
FeaturePlot(seuratM, features = "ENSG00000105976.MET", pt.size = 0.1, cols = c("lightgrey", "#BE3144"), raster = FALSE)
```

## VlnPlot MET all cell types
```{r violin MET all cell types}
###violin plot MET all cell types
seuratM$diseaseCond <- factor(seuratM$diseaseCond, levels=c("healthy", "explant", "visit1", "visit2", "visit3"))
Idents(seuratM) <- seuratM$diseaseCond
levels(seuratM)
VlnPlot(object=seuratM, features = "ENSG00000105976.MET", cols = coldiseaseCond, pt.size = 0.5, raster = FALSE) + theme(legend.position = "none")
```

## VlnPlot MET in CM
```{r violin MET CM}
###violin plot MET on CM
seuratCM <- subset(seuratM, clusterName %in% c("CM"))
table(seuratCM$clusterName)
seuratCM$diseaseCond <- factor(seuratCM$diseaseCond, levels=c("healthy", "explant", "visit1", "visit2", "visit3"))
Idents(seuratCM) <- seuratCM$diseaseCond
levels(seuratCM)
VlnPlot(object=seuratCM, features = "ENSG00000105976.MET", cols = coldiseaseCond, pt.size = 0.5, raster = FALSE) + theme(legend.position = "none")
```

## VlnPlot HGF all cell types
```{r violin HGF all cell types}
###violin plot HGF all cell types
seuratM$diseaseCond <- factor(seuratM$diseaseCond, levels=c("healthy", "explant", "visit1", "visit2", "visit3"))
Idents(seuratM) <- seuratM$diseaseCond
levels(seuratM)
VlnPlot(object=seuratM, features = "ENSG00000019991.HGF", cols = coldiseaseCond, pt.size = 0.5, raster = FALSE) + theme(legend.position = "none")
```

## VlnPlot HGF in Fb
```{r violin HGF Fb}
###violin plot HGF on Fb
seuratFb <- subset(seuratM, clusterName %in% c("Fb"))
table(seuratFb$clusterName)
seuratFb$diseaseCond <- factor(seuratFb$diseaseCond, levels=c("healthy", "explant", "visit1", "visit2", "visit3"))
Idents(seuratFb) <- seuratFb$diseaseCond
levels(seuratFb)
VlnPlot(object=seuratFb, features = "ENSG00000019991.HGF", cols = coldiseaseCond, pt.size = 0.5, raster = FALSE) + theme(legend.position = "none")
```

## VlnPlot BMP4 in Fb
```{r violin BMP4 Fb}
###violin plot BMP4 on Fb
seuratFb <- subset(seuratM, clusterName %in% c("Fb"))
table(seuratFb$clusterName)
seuratFb$diseaseCond <- factor(seuratFb$diseaseCond, levels=c("healthy", "explant", "visit1", "visit2", "visit3"))
Idents(seuratFb) <- seuratFb$diseaseCond
levels(seuratFb)
VlnPlot(object=seuratFb, features = "ENSG00000125378.BMP4", cols = coldiseaseCond, pt.size = 0.5, raster = FALSE) + theme(legend.position = "none")
```

## VlnPlot HGF in PerivFb
```{r violin HGF PerivFb}
###violin plot HGF on PerivFb
seuratPerivFb <- subset(seuratM, clusterName %in% c("PerivFb"))
table(seuratPerivFb$clusterName)
seuratPerivFb$diseaseCond <- factor(seuratPerivFb$diseaseCond, levels=c("healthy", "explant", "visit1", "visit2", "visit3"))
Idents(seuratPerivFb) <- seuratPerivFb$diseaseCond
levels(seuratPerivFb)
VlnPlot(object=seuratPerivFb, features = "ENSG00000019991.HGF", cols = coldiseaseCond, pt.size = 0.5, raster = FALSE) + theme(legend.position = "none")
```

## VlnPlot HGF in Mph
```{r violin HGF Mph}
###violin plot HGF on Mph
seuratMph <- subset(seuratM, clusterName %in% c("Mph", "Mph2"))
table(seuratMph$clusterName)
seuratMph$diseaseCond <- factor(seuratMph$diseaseCond, levels=c("healthy", "explant", "visit1", "visit2", "visit3"))
Idents(seuratMph) <- seuratMph$diseaseCond
levels(seuratMph)
VlnPlot(object=seuratMph, features = "ENSG00000019991.HGF", cols = coldiseaseCond, pt.size = 0.5, raster = FALSE) + theme(legend.position = "none")
```


## VlnPlot BMP5 in PerivFb
```{r violin BMP5 PerivFb}
###violin plot BMP5 on PerivFb
seuratPerivFb <- subset(seuratM, clusterName %in% c("PerivFb"))
table(seuratPerivFb$clusterName)
seuratPerivFb$diseaseCond <- factor(seuratPerivFb$diseaseCond, levels=c("healthy", "explant", "visit1", "visit2", "visit3"))
Idents(seuratPerivFb) <- seuratPerivFb$diseaseCond
levels(seuratPerivFb)
VlnPlot(object=seuratPerivFb, features = "ENSG00000112175.BMP5", cols = coldiseaseCond, pt.size = 0.5, raster = FALSE) + theme(legend.position = "none")
```

## BMP vs Tcells and Mph
```{r BMP4vsTcells}
###rel abundance CD8 Tcells
datList <- NULL
for(con in unique(seuratM$patient_diseaseCond)){
  seuratSub <- subset(seuratM, patient_diseaseCond==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$clusterName)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(patient_diseaseCond=con)
  datList[[con]] <- dat_con
}
dat_allclust <- do.call("rbind", datList)
## CD8 Tcells
datTcell <- dat_allclust %>% filter(Var1 == "Tcell") 
rownames(datTcell) <- datTcell$patient_diseaseCond

## Mph
datMph <- dat_allclust %>% filter(Var1 == "Mph") 
rownames(datMph) <- datMph$patient_diseaseCond

###BMP4 expr fb
Idents(seuratFb) <- seuratFb$patient_diseaseCond
levels(seuratFb)
BMP4 <- "ENSG00000125378.BMP4"

avg.all <- AverageExpression(seuratFb, group.by = "patient_diseaseCond")
avg.all <- data.frame(log1p(avg.all$RNA)) %>% 
  rownames_to_column(var="gene")
avg.BMP4 <- avg.all  %>% filter(gene == BMP4)
avg.BMP4 <- avg.BMP4 %>% tibble::column_to_rownames(var="gene")
avg.BMP4t <- data.frame(t(avg.BMP4)) 
avg.BMP4t$patient_diseaseCond <- rownames(avg.BMP4t)
avg.BMP4t <- avg.BMP4t %>% mutate(patient_diseaseCond=gsub("\\.", "_", patient_diseaseCond))

dat_all <- merge(datTcell, avg.BMP4t, by = "patient_diseaseCond")
dat_allMph <- merge(datMph, avg.BMP4t, by = "patient_diseaseCond")

dat_all <- dat_all %>% mutate(diseasecond=gsub(".*_","",patient_diseaseCond)) %>% mutate(patient=gsub("_.*","",patient_diseaseCond))
dat_allMph <- dat_allMph %>% mutate(diseasecond=gsub(".*_","",patient_diseaseCond)) %>% mutate(patient=gsub("_.*","",patient_diseaseCond))
```

## scatter BMP vs Tcells all
```{r scatter BMP4vsTcells all}
p <- ggscatter(dat_all, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient_diseaseCond",
                     color = "diseasecond",
                     palette = c("#dfc27d","#BE3144","#202547","#355C7D","#779d8d"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "TcellFreq",
                     ylab = "BMP4",
                     font.label = c("0", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0, label.y = -0.2)  # Add correlation stats
p
```

## scatter BMP vs Mph all
```{r scatter BMP4vsMph all}
p <- ggscatter(dat_allMph, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient_diseaseCond",
                     color = "diseasecond",
                     palette = c("#dfc27d","#BE3144","#202547","#355C7D","#779d8d"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "MphFreq",
                     ylab = "BMP4",
                     font.label = c("0", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.1, label.y = 0)  # Add correlation stats
p
```

## scatter BMP vs Tcells individual
```{r scatter BMP4vsTcells individual diseaecond}
dat_HH <- dat_all %>% filter(diseasecond == "healthy") 

p <- ggscatter(dat_HH, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#dfc27d"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "TcellFreq",
                     ylab = "BMP4",
                     font.label = c("10", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.020, label.y = 1)  # Add correlation stats
p


dat_expl <- dat_all %>% filter(diseasecond == "explant") 
p <- ggscatter(dat_expl, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#BE3144"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "TcellFreq",
                     ylab = "BMP4",
                     font.label = c("10", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.04, label.y = 1.5)  # Add correlation stats
p

dat_v1to3 <- dat_all %>% filter(diseasecond %in% c("visit1", "visit2", "visit3"))
p <- ggscatter(dat_v1to3, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#202547","#355C7D","#779d8d"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "TcellFreq",
                     ylab = "BMP4",
                     font.label = c("0", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.2, label.y = 0.75)  # Add correlation stats
p


dat_v1 <- dat_all %>% filter(diseasecond %in% c("visit1"))
p <- ggscatter(dat_v1, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#202547"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "TcellFreq",
                     ylab = "BMP4",
                     font.label = c("10", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.2, label.y = 0.8)  # Add correlation stats
p


dat_v2 <- dat_all %>% filter(diseasecond %in% c("visit2"))
p <- ggscatter(dat_v2, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#355C7D"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "TcellFreq",
                     ylab = "BMP4",
                     font.label = c("10", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.1, label.y = 0.1)  # Add correlation stats
p


dat_v3 <- dat_all %>% filter(diseasecond %in% c("visit3"))
p <- ggscatter(dat_v3, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#779d8d"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "TcellFreq",
                     ylab = "BMP4",
                     font.label = c("10", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.1, label.y = 0.7)  # Add correlation stats
p
```

## scatter BMP vs Mph individual
```{r scatter BMP4vsMph individual diseaecond}
dat_HH <- dat_allMph %>% filter(diseasecond == "healthy") 

p <- ggscatter(dat_HH, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#dfc27d"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "MphFreq",
                     ylab = "BMP4",
                     font.label = c("10", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.1, label.y = 0.1)  # Add correlation stats
p


dat_expl <- dat_allMph %>% filter(diseasecond == "explant") 
p <- ggscatter(dat_expl, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#BE3144"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "MphFreq",
                     ylab = "BMP4",
                     font.label = c("10", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.1, label.y = 1.5)  # Add correlation stats
p

dat_v1to3 <- dat_allMph %>% filter(diseasecond %in% c("visit1", "visit2", "visit3"))
p <- ggscatter(dat_v1to3, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#202547","#355C7D","#779d8d"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "MphFreq",
                     ylab = "BMP4",
                     font.label = c("0", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.2, label.y = 0.75)  # Add correlation stats
p


dat_v1 <- dat_allMph %>% filter(diseasecond %in% c("visit1"))
p <- ggscatter(dat_v1, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#202547"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "MphFreq",
                     ylab = "BMP4",
                     font.label = c("10", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.15, label.y = 0.8)  # Add correlation stats
p


dat_v2 <- dat_allMph %>% filter(diseasecond %in% c("visit2"))
p <- ggscatter(dat_v2, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#355C7D"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "MphFreq",
                     ylab = "BMP4",
                     font.label = c("10", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.1, label.y = 0.1)  # Add correlation stats
p


dat_v3 <- dat_allMph %>% filter(diseasecond %in% c("visit3"))
p <- ggscatter(dat_v3, x="percent", y="ENSG00000125378.BMP4",
                     label= "patient",
                     color = "diseasecond",
                     palette = c("#779d8d"),
                     repel = T,
                     label.rectangle = F,
                     xlab = "MphFreq",
                     ylab = "BMP4",
                     font.label = c("10", "plain","black"),
                     size=2,) +
  geom_point(aes(color = diseasecond), alpha = 0.5, size = 3) +  # Customize point size and transparency
  geom_smooth(method = "lm", se = TRUE, color = "black") + # Add regression line with confidence interval
  stat_cor(aes(label = paste(..r.label.., ..rr.label.., ..p.label.., sep = "~`,`~")), label.x = 0.2, label.y = 0.7)  # Add correlation stats
p
```

## session info
```{r date and session info}
date()
sessionInfo()
```
