---
title: "IntCellChat"
author: "A.DeMartin"
date: "2024-10-04"
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

## load packages 
```{r load packages}
library(Seurat)
```

## load packages cellchat
```{r load packages cellchat}
#devtools::install_github("sqjin/CellChat")
library(CellChat)
library(patchwork)
options(stringsAsFactors = FALSE)
```

## load file
```{r load merged file, eval=FALSE, include=TRUE}
##load merged file 
fileNam <- "/Users/immbio/Desktop/Project/Angelina/HumanHeartCarTrans/data/Human_heart_allmerged_seurat.rds"
seuratM <- readRDS(fileNam)
table(seuratM$orig.ident)
```

## assign clusters
```{r assign, eval=FALSE, include=TRUE}
#### clusterName
## to get this order: levels=c("CM","Fb","PerivFb","VSMC","Int","BEC1", "BEC2","LEC","NC","AdipoC","Mph","Mph2","Tcell","Tcell2")
seuratM$clusterName <- "clusterName"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "0" )] <- "b_Fb"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "1" )] <- "f_BEC1"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "2" )] <- "c_PerivFb"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "3" )] <- "k_Mph"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "4" )] <- "a_CM"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "5" )] <- "m_Tcell"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "6" )] <- "d_VSMC"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "7" )] <- "l_Mph2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "8" )] <- "g_BEC2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "9" )] <- "e_Int"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "10" )] <- "i_NC"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "11" )] <- "n_Tcell2"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "12" )] <- "h_LEC"
seuratM$clusterName[which(seuratM$RNA_snn_res.0.25 %in% "13" )] <- "j_AdipoC"
table(seuratM$clusterName)
table(seuratM$RNA_snn_res.0.25)
```

convert to sce to change rownames
```{r make sce,  eval=FALSE, include=TRUE}
sce <- as.SingleCellExperiment(seuratM)
#scecko <- as.SingleCellExperiment(seuratcko)
#scewt <- as.SingleCellExperiment(seuratwt)
#rownames(sce) = gsub("\\..*","",rownames(sce))
rownames(sce) = gsub("^.*\\.","",rownames(sce))
#rownames(scecko) = gsub("^.*\\.","",rownames(scecko))
#rownames(scewt) = gsub("^.*\\.","",rownames(scewt))
```

create a CellChat object
```{r create CellChat object,  eval=FALSE, include=TRUE}
cellchat <- createCellChat(object = sce, group.by = "clusterName")
#cellchat <- createCellChat(object = scecko, group.by = "clusterLabelext")
#cellchat <- createCellChat(object = scewt, group.by = "clusterLabelext")
#> Create a CellChat object from a data matrix
#> Set cell identities for the new CellChat object
```

set the ligand receptor interaction database
```{r set database,  eval=FALSE, include=TRUE}
CellChatDB <- CellChatDB.human # use CellChatDB.mouse if running on mouse data
showDatabaseCategory(CellChatDB)
# Show the structure of the database
dplyr::glimpse(CellChatDB$interaction)

# use a subset of CellChatDB for cell-cell communication analysis
#CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling") # use Secreted Signaling

# use all CellChatDB for cell-cell communication analysis
CellChatDB.use <- CellChatDB # simply use the default CellChatDB

# set the used database in the object
cellchat@DB <- CellChatDB.use
```

processing expression data for cellcell communication analysis
```{r processing expression data,  eval=FALSE, include=TRUE}
# subset the expression data of signaling genes for saving computation cost
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
future::plan("multisession", workers = 4) # do parallel
options(future.globals.maxSize = 1000 * 1024^2)
#> Warning: [ONE-TIME WARNING] Forked processing ('multicore') is disabled
#> in future (>= 1.13.0) when running R from RStudio, because it is
#> considered unstable. Because of this, plan("multicore") will fall
#> back to plan("sequential"), and plan("multiprocess") will fall back to
#> plan("multisession") - not plan("multicore") as in the past. For more details,
#> how to control forked processing or not, and how to silence this warning in
#> future R sessions, see ?future::supportsMulticore
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
# project gene expression data onto PPI network (optional)
#cellchat <- projectData(cellchat, PPI.human)
```

Compute the communication probability and infer cellular communication network
```{r compute communication probability, eval=FALSE, include=TRUE}
cellchat <- computeCommunProb(cellchat, type =  "triMean")
cellchatH <- computeCommunProb(cellchat, type =  "truncatedMean", trim= 0.1) #set truncated mean to 10% - average gene expression is zero if less than 10% of cell in one group express gene (default is 25%)
# Filter out the cell-cell communication if there are only few number of cells in certain cell groups
# cellchat <- filterCommunication(cellchat, min.cells = 35)
```

Infer the cell-cell communication at a signaling pathway level
CellChat computes the communication probability on signaling pathway level by summarizing the communication probabilities of all ligands-receptors interactions associated with each signaling pathway.
```{r CommunProbPathwqy, eval=FALSE, include=TRUE}
cellchat <- computeCommunProbPathway(cellchat)
cellchatH <- computeCommunProbPathway(cellchatH)
```

```{r aggregateNet, eval=FALSE, include=TRUE}
cellchat <- aggregateNet(cellchat)
cellchatH <- aggregateNet(cellchatH)
```

Compute the network centrality scores
```{r computeCentrality, eval=FALSE, include=TRUE}
future.seed=TRUE
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP") # the slot 'netP' means the inferred intercellular communication network of signaling pathways
cellchatH <- netAnalysis_computeCentrality(cellchatH, slot.name = "netP") # the slot 'netP' means the inferred intercellular communication network of signaling pathways
```


```{r save CellChat object, eval=FALSE, include=TRUE}
saveRDS(cellchat, file = "/Users/immbio/Desktop/Project/Angelina/HumanHeartCarTrans/Human_heart_allmerged_cellchat.rds")
saveRDS(cellchatH, file = "/Users/immbio/Desktop/Project/Angelina/HumanHeartCarTrans/Human_heart_allmerged_cellchatH.rds")
```

############################################################# end pre-processing ########################################################

```{r load cellchat object}
cellchat <- readRDS("/Users/immbio/Desktop/Project/Angelina/HumanHeartCarTrans/data/Human_heart_allmerged_cellchat.rds")
cellchatH <- readRDS("/Users/immbio/Desktop/Project/Angelina/HumanHeartCarTrans/data/Human_heart_allmerged_cellchatH.rds")
```

```{r setcolor vector}
colclusterName <- c("#67001f", "#D53E4F", "#f4a582", "#FEE08B", "#003c30","#01665e","#66C2A5", "#3288BD","#BEAED4", "#c7eae5","#355C7D","#202547","#B45B5C","#8c510a")
```

Extract the inferred cellular communication network as a data frame
We provide a function subsetCommunication to easily access the inferred cell-cell communications of interest
```{r extract interacitons}
df.netH <- subsetCommunication(cellchatH)
df.net <- subsetCommunication(cellchat) #returns a data frame consisting of all the inferred cell-cell communications at the level of ligands/receptors. Set slot.name = "netP" to access the the inferred communications at the level of signaling pathways

#interactions immunecells to stroma only
#df.net <- subsetCommunication(cellchat.all) 
#sources.use = c(1:11), targets.use = c(12:20)

#df.net <- subsetCommunication(cellchat, signaling = c("LAMININ")) #gives the inferred cell-cell communications mediated by signaling LAMININ
```

Calculate the aggregated cell-cell communication network
We can calculate the aggregated cell-cell communication network by counting the number of links or summarizing the communication probability. USER can also calculate the aggregated network among a subset of cell groups by setting sources.use and targets.use.
```{r all signaing}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
gg1 <- netAnalysis_signalingRole_scatter(cellchat, color.use = colclusterName)
gg1
gg1 <- netAnalysis_signalingRole_scatter(cellchatH, color.use = colclusterName)
gg1
```


```{r all signaing heatmap}
# Heatmap
par(mfrow=c(1,1))
netVisual_heatmap(cellchat, color.heatmap = "GnBu", color.use = colclusterName, measure = "count")
netVisual_heatmap(cellchat, color.heatmap = "GnBu", color.use = colclusterName, measure = "weight")

par(mfrow=c(1,1))
netVisual_heatmap(cellchatH, color.heatmap = "GnBu", color.use = colclusterName, measure = "count")
netVisual_heatmap(cellchatH, color.heatmap = "GnBu", color.use = colclusterName, measure = "weight")
```
CellChat can also visualize the aggregated cell-cell communication network. For example, showing the number of interactions or the total interaction strength (weights) between any two cell groups using circle plot.
```{r number of int btw celltypes}

groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")

groupSize <- as.numeric(table(cellchatH@idents))
netVisual_circle(cellchatH@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchatH@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

```{r number of int btw celltypes indiv, fig.height=10, fig.width=10}
mat <- cellchat@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}

mat <- cellchatH@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```

```{r heatmap pathways, fig.height= 10, fig.width= 10}
#> Do heatmap based on a single object
#Identify signals contributing most to outgoing or incoming signaling of certain cell groups
#We can also answer the question on which signals contributing most to outgoing or incoming signaling of certain cell groups.
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", color.use = colclusterName)
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", color.use = colclusterName)
ht1 + ht2

ht1 <- netAnalysis_signalingRole_heatmap(cellchatH, pattern = "outgoing", color.use = colclusterName, height = 15)
ht2 <- netAnalysis_signalingRole_heatmap(cellchatH, pattern = "incoming", color.use = colclusterName, height = 15)
ht1 + ht2
```

```{r sender/receiver}
# Visualize the computed centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups
pathways.show <- c("LAMININ")
netAnalysis_signalingRole_network(cellchatH, signaling = pathways.show, width = 8, height = 2.5, font.size = 10, color.heatmap = "Blues", color.use = colclusterName)

pathways.show <- c("COLLAGEN")
netAnalysis_signalingRole_network(cellchatH, signaling = pathways.show, width = 8, height = 2.5, font.size = 10, color.heatmap = "Blues", color.use = colclusterName)

pathways.show <- c("FN1")
netAnalysis_signalingRole_network(cellchatH, signaling = pathways.show, width = 8, height = 2.5, font.size = 10, color.heatmap = "Blues", color.use = colclusterName)
```

```{r circle plot COLLAGEN}
pathways.show <- c("COLLAGEN")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[2,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```
```{r circle plot CXCL}
pathways.show <- c("CXCL")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[2,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```
```{r circle plot CCL}
pathways.show <- c("CCL")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[2,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```

```{r circle plot FGF}
pathways.show <- c("FGF")

pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,]
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[2,] 
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[5,]
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[6,] 
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```

```{r circle plot NCAM}
pathways.show <- c("NCAM")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[2,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[3,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```
```{r circle plot BMP}
pathways.show <- c("BMP")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[2,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[3,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```
```{r circle plot SEMA5}
pathways.show <- c("SEMA5")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[2,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[3,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[4,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```
```{r circle plot TGFB}
pathways.show <- c("TGFb")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[2,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[3,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[4,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```

```{r circle plot CDH}
pathways.show <- c("CDH")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```
```{r circle plot EGF}
pathways.show <- c("EGF")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```
```{r circle plot NEGR}
pathways.show <- c("NEGR")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```
```{r circle plot COMPLEMENT}
pathways.show <- c("COMPLEMENT")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[2,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[3,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```
```{r circle plot TENASCIN}
pathways.show <- c("TENASCIN")
pairLR <- extractEnrichedLR(cellchatH, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR[1,] # first in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[2,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)

LR.show <- pairLR[3,] # second in list shown
## Circle plot
netVisual_individual(cellchatH, signaling = pathways.show, pairLR.use = LR.show,layout = "circle", color.use = colclusterName, weight.scale = T)
```

## session info
```{r date and session info}
date()
sessionInfo()
```